---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.14.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Lab 7


## Estimating a Probability via Simulation -- Example 1

In this part of the lab, we will use computer simulation to estimate a probability.

Suppose there are two boxes, A and B.  Box A contains tickets {1,2,3,4,5} and box B contains tickets {1,3,5}.  Let $X$ be a random draw from $A$ and $Y$ be a random draw from B.  Let $Z = X + Y$.  What is the probability that $Z$ is equal to 5, 6, or 7?  

We will answer this question in two ways: (1) Exactly, by enumerating all possible outcomes; and (2) by computer simulation.
```{r}
a = c(1, 1, 2, 4, 5)
sample(a, 1, replace = TRUE)
```



### Exact calculation

Calculate the exact probability that $Z$ is equal to 5, 6, or 7.  Hint: there are a total of $5\times 3 = 15$ possible outcomes.  


**Answer**:

```{r}
## Write your code here
record = c()
A = c(1,2,3,4,5)
B = c(1,3,5)
for (i in A) {
  for (j in B) {
    record = c(record, i+j)
  }
}
sum(record == 5 | record == 6 | record == 7)
```

### Computer Simulation

You will now write code to find the probability using computer simulation.


#### Step 1: Simple simulation

Using the `sample` command, write code that generates a random number X, a random number Y, and the sum Z.  Then print Z.  

```{r}
## Write your code here
  X = sample(A,1)
  Y = sample(B,1)
  Z = X + Y
  print (Z)
```

#### Step 2: Repeating the simulation

Now create a for loop that runs the above code 100 times (and prints Z each time).  (Note: you will want to copy and paste your above code into a for loop.)

```{r}
## Write your code here
for (i in 1:100) {
  X = sample(A,1)
  Y = sample(B,1)
  Z = X + Y ##Zs are independent
  print (Z)
}
```

#### Step 3: Counting the times Z equals 5, 6, or 7.

Copy and paste your above code but remove the print function and increase the number of repetitions in the for loop from 100 to 10000.  Then, add a few lines that will count the number of times Z equals 5, 6, or 7.  Specifically, as the very first line, create a variable called count and set it equal to zero.  Then, inside the for loop, add an `if` statement that checks whether Z is equal to 5, 6, or 7 and, if so, increments the count variable by 1.

Then, at the end of the code, create a variable called `estimated_probability` that is equal to `count / 10000`, and then print it. 

```{r}
## Write your code here
count = 0
for (i in 1:10000) {
  X = sample(A, 1)
  Y = sample(B,1)
  Z = X + Y
  if (Z == 5 | Z == 6 | Z == 7) {
    count = count+1
  }
}
count/10000
```

## Estimating a Probability via Simulation -- Example 2

This part of the lab is similar to the preceding part, although the example is slightly different, and you will complete it on your own.

Suppose there are two boxes, A and B.  Box A contains tickets {-2, -1, 0, 1, 2}, box B contains tickets {-1,2}, and box C contains tickets {0, 1}.  Let $X$ be a random draw from $A$ and $Y$ be a random draw from B and $Z$ be a random draw from C.  Let $U = XY + Z$.  What is the probability that $U$ is equal to 0?  

We will answer this question in two ways: (1) Exactly, by enumerating all possible outcomes; and (2) by computer simulation.


### Exact calculation

Calculate the exact probability that $U$ is equal to 0.  

```{r}
## Write your code here
A = c(-2, -1, 0, 1,2)
B = c(-1, 2)
C = c(0,1)
record = c()
for (i in A) {
  for (j in B) {
    for (k in C) {
      record = c(record, i + j + k)
    }
  }
}
print (record)
```


**Answer**:

<< Type your answer here >>


### Computer Simulation

You will now write code to find the probability using computer simulation.


#### Step 1: Simple simulation

Using the `sample` command, write code that generates a random number X, a random number Y, a random number Z, and the random number U.  Then print U.  

```{r}
## Write your code here
count = 0
for (i in 1:10000) {
  X = sample(A, 1)
  Y = sample(B,1)
  Z = sample(C, 1)
  if (Z == X + Y) {
    count = count+1
  }
}
count/10000
```


#### Step 2: Repeating the simulation

Now create a for loop that runs the above code 100 times (and prints U each time).  

```{r}
## Write your code here

count = 0
for (i in 1:100) {
  X = sample(A, 1)
  Y = sample(B,1)
  Z = sample(C, 1)
  if (Z == X + Y) {
    count = count+1
  }
}
count/100
```

#### Step 3: Counting the times Z equals 5, 6, or 7.

Copy and paste your above code but remove the print function and increase the number of repetitions in the for loop from 100 to 10000.  Then, add a few lines that will count the number of times U equals 0, and divide this by 10000 to estimate the probability that U equals 0.  

## Review Exercises

**These do not need to be completed as part of the lab.**

The review exercises are for additional practice.  They may appear on exams, or on homework assignments.  But it is not necessary to complete them as part of the lab assignment.


1\. Consider the following code.  What is the probability the output is TRUE?

```{r}
box1 = c(3,5,7)
box2 = c(-1,1)
x = sample(box1, 1)
y = sample(box2, 1)
print(x+y == 6)
```

2\. Consider the following code.  The output will be approximately equal to what number?

```{r}
box1 = c(3,5,7)
box2 = c(-1,1)
count = 0
for (i in 1:10000)
    {
    x = sample(box1, 1)
    y = sample(box2, 1)
    if (x+y == 6) count = count + 1
}
print(count / 10000)
```

3\. Consider again the code in problem 2.  Every time you run the code, you will get a slightly different answer.  Typically, about how far off will the output of the code be from the answer you gave for problem 2?

```{r}
## Write your code here
prob = c()
for (i in 1:100) {
count = 0
for (i in 1:100) {
    x = sample(box1, 1)
    y = sample(box2, 1)
    if (x+y == 6) count = count + 1
}
 prob[i] = count / 100
}
 prob
```

```{r}
## Write your code here
var <- function(x) {
return (mean((x-mean(x))^2))
}
var(prob)
```

4\. Consider the following code.  The output will be approximately equal to what number?

```{r}
box1 = c(3,5,7)
box2 = c(-1,0,1)
count = 0
for (i in 1:10000)
    {
    x = sample(box1, 1)
    y = sample(box2, 1)
    if (x*y < 4) count = count + 1
}
print(count / 10000)
```

5\. Consider again the code in problem 4.  Every time you run the code, you will get a slightly different answer.  Typically, about how far off will the output of the code be from the answer you gave for problem 4?


6\. Consider the following code.  The output will be approximately equal to what number?

```{r}
box1 = c(3,5,7)
box2 = c(-1,1)
not_really_a_count = 0
for (i in 1:10000)
    {
    x = sample(box1, 1)
    y = sample(box2, 1)
    if (x+y == 6) not_really_a_count = not_really_a_count + 1
    if (x+y == 8) not_really_a_count = not_really_a_count + 2
}
print(not_really_a_count / 10000)
```

```{r}
qnorm(0.001, 0, 1, lower.tail=FALSE)
```

```{r}
print(die_rolls)
```

