---
title: "STATS 280 Lab 9 Solutions"
author: "Jianfei Wang and Vincenzo Loffredo"
date: "2023-03-22"
output: html_document
---


# Confidence Intervals

Confidence intervals are a statistic (function of the data) used to gauge uncertainty about an estimate.
In frequentist statistics the assumption is that there exists a true parameter $\mu$ that we are trying to estimate using the data $X$ and a statistics $T(X)$, like the mean or the standard deviation of the data.

A confidence interval is an interval that contains the population parameter with a certain level of confidence ($\alpha$), and generally takes the form:
\begin{align}
CI_{\alpha}(X) & = \text{(point estimate)} +/- \text{(critical value)*(standard error)}.
\end{align}
Therefore a confidence interval is a function of the sample, and for each sample we will get a different CI.

Whenever we can use the Normal approximation (Central Limit Theorem and adjecent results), we can use the  critical value formulation:

\begin{align}
CI_{\alpha}(X) & = \text{(point estimate)} +/- z_{\alpha/2}*\text{(standard error)},
\end{align}

where $z_{\alpha/2}$ is the critical value from the standard normal distribution.


## Box Model Applications

### Theory Exercise

Consider a sample of size $1600$ from a box.
The mean of the sample is $100$ and its standard deviation is 20.
Build a $75\%$ confidence interval for the mean of the box.

### Coding

```{r}
# hint: to get the critical value of for 95% confidence interval with normal approximation
# crit = qnorm(0.05/2, 0, 1, lower.tail = F)

```

### Solutions
```{r}
crit = qnorm( (1-0.75)/2, 0, 1, lower.tail = F)
n_mean = 100
ste = 20/sqrt(1600)
CI2 = c(n_mean - crit*ste, n_mean + crit*ste)
```

### Theory Exercise

Suppose a $0-1$ box contains $2000$ tickets. 
A researcher samples $200$ tickets from the box with replacement and the sample has a mean of $0.75$ and a standard deviation of $0.4341$.
What is the $95\%$ confidence interval for mean of the box?

### Coding Exercise

Build a simulation to take a sample of size $200$ with replacement from the box and compute the confidence interval for the mean.
Suppose the $0-1$ box contains $500$ '0' tickets and $1500$ '1'. 

```{r}
# hint: the rep(a,b) command outputs a vector which repeats a for b number of times.

# build up the box
box = c (rep(0, 500), rep(1, 1500))
s = sample (box, 200, replace = T) 
s_mean = mean(s)
s_sd = sd(s)
ste = s_sd / sqrt(200)
CI = c(s_mean - 2*ste, s_mean + 2*ste)

# sample from the box
# s = sample (box, 200, replace = T) 
 
# Compute the sample mean and the sample standard deviation 
# s_mean =  mean(s)
# s_sd = sd(s)
#ste = s_sd / sqrt(200)
  
# build up the 95% confidence interval with the sample mean and standard error
# CI = 
```

### Solutions

```{r}
set.seed(1)
box = c(rep(0,500),rep(1,1500))
s = sample(box, 200, replace = F)
s_mean = mean(s)
s_sd = sd(s)
CI = c(s_mean - 2*ste, s_mean + 2*ste)
```



## Using Linear Regression for Confidence Intervals

We use a built-in dataset called mtcars, with data from cars in the 70s.

```{r}
# import the mtcars dataset
data("mtcars")
summary(mtcars)

```

We have seen how to run a linear model for simple linear regression in the previous labs, and now we are going to show how to  use the `lm()` function to compute confidence intervals for a specific variable.

If we want to compute the confidence interval for the mean of one variable, say the miles per gallon, we can write

```{r}
mod = lm(formula = mpg ~1, data = mtcars)
summary(mod)
```

In this code:

- lm is the function used to create a linear model.
- The argument data is used to specify the dataset from which the variable names (mpg in our case) are taken from.
- Formula gives the formula of the model: the first variable is the variable of interest (Y), while the ones on the other side of the `~` are the covariates used in the model.
- Using only $1$ as a covariate means to only have an intercept, this means to just estimate the mean of the variable.


The most straightforward way to get a confidence interval for the coefficients of a linear model is `confint()`, a function that takes as arguments a linear model and a level $\alpha$. 
In our case, we have only the intercept, so the coefficient for the intercept is going to be the mean of the variable.

If not specified, $\alpha=0.05$, so the confidence interval will be at $95\%$.

```{r}
confint(object = mod , level = 0.95)
```

This interval is going to be slightly different than the one obtained using the Normal approximation.
To see the difference, we can calculate by hand the confidence interval with normal approximation as follows:

```{r}
mean = mean(mtcars$mpg)
n = nrow(mtcars)
sd = sd(mtcars$mpg) #Don't worry here about the distinction between sd as calculated by R and sd as calculated in class.
se = sd(mtcars$mpg)/sqrt(n)
print(c(mean - 2*se,mean+2*se))
```

The confidence interval results are different since R uses Student's t distribution to get the critical value.

Long story short, when having low sample size, the sample standard deviation as the estimate of population standard deviation is adds more uncertainty. 

The t-distribution is basically a Normal distribution with heavier tails to  account for this additional uncertainty.
The way we build a confidence interval stays the same, but we don't use the Normal distribution for critical values, but the T distribution.

As the sample size increases, the t distribution approaches a standard normal distribution, and so do the critical values, making the confidence interval grow closer.
We will learn more about T-distributions in the following labs, but this was just a sneak peak.

We can use this model formulation also for our previous box:

```{r}
mod = lm(s ~1)
# R - using T distribution
confint(mod)
# Manual (normal approximation)
es = coef(summary(mod))[1,1]
se = coef(summary(mod))[1,2]
c(es - 2*se, es+2*se)
```
As you can see, with a higher sample size, the two confidence intervals are much closer.


## Coverage proability

Coverage is the percentage of time that the confidence interval contains the true value of the population parameter. 
If we construct 95% confidence intervals from many random samples, the coverage should be roughly 95%. 
Sometimes confidence intervals could be calculated with different methods, we can use coverage to review their performances. 

### Coding Exercise 

Suppose we still have the 01 box that contains 2000 tickets, 500 of which are 0 and 1500 of which are 1. We sample 200 tickets. What is the coverage of confidence intervals by using normal approximation if we repeat the simulations for 1000 times?


```{r}
# hint: run the above simulations for 1000 times, create lower bound and upper bound vectors of confidence interval for 1000 iterations.


# build up the box
box = c(rep(0, 500), rep (1, 1500))
s = sample(box, 200, replace = T)
# Compute mean of the box
mu = mean(box)

# Create a storage vector for how many times the mean is in the confidence interval
count = 0

# Start the loop
for(i in 1:1000){
s = sample(box, 200, replace = TRUE)
avg = mean(s)
stdv = sd(s)
ste = stdv/sqrt(200)
lower = avg - 2 * ste
upper = avg + 2 * ste
if (lower <= mu & mu <= upper) {
  count = count + 1
}
}
print (count/1000)

```


### Solutions (Alternative approach)
```{r}
set.seed(1)
mu= mean(box)
lower = c()
upper = c()
for (i in 1:1000){
  samples = sample(box, 200, replace = F)
  m = mean(samples)
  s = sd(samples)
  ste = sqrt((2000-200)/(2000-1)) * s / sqrt(200)
  lower[i] = m - 2*ste
  upper[i] = m + 2*ste
}
mean(lower <= mu & mu <= upper)
```


```{r}
data("mtcars")
```

```{r}
m = mean(mtcars$wt)
print(m)
sd = sd(mtcars$wt)
se = sd/sqrt((length(mtcars$wt)))
z = 1.645
ci = c(m-se*z, m+se*z)
print (ci)
```

```{r}

```

